JD notes on effects

### 2022-06-14 manuscript/body.tex
Intro changed

Ordinary least squares (OLS) regression provides an easy way to express multivariate relationships as an additive linear combination of many predictors with the outcome. The coefficient estimates from OLS regression models are generally easy to interpret as the expected change in the outcome with a unit change in the continuous predictor or as the expected outcome difference between the levels of the categorical predictor with the baseline category. However, with additional complexities such as interactions between predictors, the resulting coefficients risk misinterpretation or interpretation complexity \citep{brambor_understanding_2006, berry_improving_2012}.

Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including nonlinear relationships between response and predictors e.g., nonlinear transformation on the predictors, interactions between predictors (and via splines for example), and nonlinear transformations via link functions. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLM models involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore \emph{predictions} and \emph{effects} \citep{fox2009effect, leeper2017package, lenth2018package}.  For example, as a way to plan, public health officials may want to know the effects of household \code{income}, \code{wealthindex}, etc, on the predicted probability that slum dwellers will have improved water services.

Both predictions and effects tell us what we expect the response to be for a given value of \emph{focal} predictor (predictor of interest). The major difference between the two lies in the construction of the associated confidence intervals. In particular, for predictions, we want to use the classic curved confidence intervals (or prediction intervals) which capture the uncertainty in the whole model and generally focus on the \emph{total} effects, and so will often be suited for a univariate model. On the other hand, effects, attempt to show the effect of a focal predictor. In this case, we suggest the narrower confidence intervals which focus on uncertainty in the effect of the focal predictor only, and so will often be suited for the direct effects – after controlling for other predictors – from a multivariate model.

When visually presented, predictions and effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the predictors whose estimates are usually, but not always, harder to interpret. Generating these quantities together with the associated confidence intervals for regression models has a number of challenges. In particular:
\begin{enumerate}
\item choice of representative values of focal predictor and the \emph{reference point} for non-focal predictors, especially in multivariate models
\item  uncertainty estimation -- appropriate choice of \emph{anchor} for computing confidence intervals; and how to incorporate the uncertainty due to non-focal predictors 
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)Ms)
\end{enumerate}

The most common way of choosing the representative values is taking unique levels of the focal predictor if discrete or taking appropriately sized quantiles if continuous, and then calculating the predictions while holding non-focal (other predictors other than the focal predictor) predictors at their reference point (e.g., means) \citep{fox2009effect, hanmer2013behind}. This procedure generates -- predictions (at least in our case), \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to these quantities as predictions and effects since they should, for example, tell us what we would expect the response to be at a particular value or level of the predictor, for an ``average case''. More specifically, the basic idea is to compute the expected outcome by holding the non-focal predictors constant, for example at their mean (centered), while varying the focal predictor, with the goal that estimates represent how the model responds to the changes in the focal predictor. As mentioned above, in our case, the key difference between predictions and effects is in the computation of the confidence intervals.

The most commonly used \proglang{R} software packages for generating predictions (\pkg{emmeans} and \pkg{effects}), by default, use the average of \emph{input} and \emph{model variables} as the reference point for continuous and categorical predictors, respectively. However, there are a number of choices one can make when considering this approach -- for example, in the presence of interaction, averaging the interactions (averaging the product of interacting model variables) versus taking the product of the averages of the interacting model variables (the default for \pkg{emmeans} and \pkg{effects}). For simple models (with no interactions), the two approaches give the same predictions. However, for complex models, we demonstrate that the two approaches can produce substantially different results and show that averaging the interactions (model variables) closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal predictor:

@@

The horizontal blue, green and black lines are the mean predictions, i.e., $\bar{\hat{y}}$, the red ones are the means of the simulated $y$, i.e., $\bar{y}$, while the vertical dotted grey lines are the means of the corresponding focal predictors. The grey points are the binned simulated $y$. The trend lines represent the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other predictors at their average for the three methods and the true predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated value in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches closely match the true predictions. B: Even with the simple interaction between non-focal predictors, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and true predictions, respectively; in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not the proposed \pkg{varpred}.}

@@

In the absence of interaction (model~\ref{eq:simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, Fig~\ref{fig:justify_plots}A. However, in the presence of interaction, even as simple as the one in model~\ref{eq:simple_inter_higher}, the estimates starts to differ. In particular, \pkg{emmeans} and \pkg{effects} give similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), Fig~\ref{fig:justify_plots}B. To generate Fig~\ref{fig:justify_plots}A, all the three packages average $x_2$ and $x_3$. On other hand, to generate Fig~\ref{fig:justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction term ($x_2x_3$). In particular \pkg{emmeans} and \pkg{effects} use \emph{input variables} to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses \emph{model variables} to compute $\over{x_2x_3}$. In the case of Fig~\ref{fig:justify_plots}A, averaging the input variables is equivalent to averaging the model variables.

One may be interested in the uncertainties associated with the focal predictor of interest only, excluding other uncertainties due to other non-focal predictors -- effects. However, currently, the two packages do not provide straightforward way do achieve this. To illustrate this, we generate predictions of $x_2$ from model~\ref{eq:simple_inter_higher_no_interaction} together with the associated $95\%$ confidence bands, as shown in Fig~\ref{fig:justify_ci_plots}.

\begin{figure}[h]
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal predictor, i.e., the model center, corresponds to the effects of the focal predictor from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal predictors. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal predictor.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence bands are much wider because they include uncertainties associated with the intercept and non-focal predictors, but narrower and crosses at the mean of the focal predictor, i.e., \emph{center point or mean-based}, in \pkg{varpred}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal predictor we are more certain about, i.e., mean of the model variable corresponding to the focal predictor (center point or mean-based). For simple models, the point where the confidence bands crosses is the model center; it corresponds to anchor we choose, in this case, the mean of the focal predictor.

For linear models (including models with complex interactions), the averaging is done on the linear scale, i.e., linear averaging. As a result the bias in the expected (mean) predictions can easily be corrected by performing predictions from the \emph{model center} (average of the model variables) or the center point. The mean prediction is the prediction at the model center. On the other hand, when dealing with GL(M)Ms with nonlinear link functions, the averaging is done on the nonlinear scale, i.e., nonlinear averaging which makes estimating correct predictions much harder. In particular, bias in expected mean prediction induced by nonlinear transformation of the response variable can lead inaccurate predictions. One way to deal with this is to make predictions on the transformed scale (linear predictor scale), and then back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation, for example, second order Taylor approximation implemented in \pkg{emmeans} \citep{lenth2018package}. In linear models, mean-based (alias center point or model center) gives estimates consistent with the observed values, however, in models with nonlinear link functions, this is not always the case. An alternative to mean-based reference point, i.e., averaging non-focal predictors, is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictors and then averaging across the values of the focal predictor \citep{hanmer2013behind}.

@@

\item \textbf{Predictors}: Refer to the X-variables which go into the regression model. For example, consider the model the includes the interaction in Equation~\ref{eq:simple_inter_higher}. This regression model has $5$ predictors -- $x_1, x_2, x_3, x_2x_3$ and the intercept term.
\item \textbf{Focal and non-focal predictors}: Focal predictor refers to the predictor we are interested in generating its effects on the predictions. For multivariate models, all other remaining predictors are non-focal predictors. For example, consider model described by Equation~\ref{eq:simple_inter_higher_no_interaction}. Suppose we are interested in the effect of $x_1$ on the predicted $y$, then $x_1$ is the focal predictor while $x_2$ and $x_3$ are non-focal predictors.
\item \textbf{Input variables}: Refers to the observed (or scientific) variables underlying an inference or exploration, and forms the X-variables. For example, the regression model described by Equation~\ref{eq:simple_inter_higher} has only $4$ input variables -- $x_1, x_2, x_3$.
\item \textbf{Model matrix}: Refers to the design matrix whose rows include all combination of variables appearing in the interaction terms. Consider an example for three hypothetical households -- the first household head is a Christian with an income of $\$ 50$, the second household head is Muslim with an income of $\$ 100$ while the third household head is a Jew with an income of $\$ 77$. Suppose we want to model the household size (\code{hhsize}) as function of these household characteristics, i.e., $$\mathrm{hhsize} = \beta_0 + \beta_1\times\mathrm{income} + \beta_{2[r]}\times\mathrm{religion} + \mathrm{error}.$$ The model matrix corresponding to this model is given by

@@

\end{bmatrix}.$$ The first column represents the constant term in our model, $\beta_0$. For continuous variables, the representation in the model matrix is the same as the correspond input variables (for example second column representing income). For categorical variables, however, by default, the model matrix  creates additional dummy variables using the reference cell parameterization. This means that, if an input variable has $L$ factor levels, then there will be $L-1$ dummy columns representing all but the first level created in the model matrix. In our example, the column \code{religion} is missing and instead we have \code{religionJew} and \code{religionMuslim}; the missing category \code{religionChristian} is treated as the reference category.
\item \textbf{Model variables:} It is easy to interchangeably refer to predictors as input variables, and vice versa. To avoid this issue, we adopt the term model variables which are variables that represent columns in the model matrix. Each input variable may correspond to one or more model variables. In particular, variables with more than two categories (for instance religion in our previous example), or variables modeled with a spline or polynomial term, will correspond to more than one model variables.
\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. We refer to the prediction approach which uses model center as mean-based. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots}. In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for center point.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal predictors, when estimating the predictions. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect confidence intervals (anchor choice does not affect the main prediction estimates). Typically chosen as the center point of the model variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate CIs becomes important. This is not true for the more conventional prediction plots: we can choose any anchor, e.g., zero-anchored, but model center seems like a natural, stable choice, i.e., center-anchored. In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the minimum and center (mean) of the focal predictor.

@@

If we have more than one model variables for our focal predictor, however, the model center does not correspond to any given value of the variable, and we don't get the nice crossing we get in simpler cases. See Fig~\ref{fig:pred_cubic_plots}.

\subsection{Prediction- vs. effect- styled plots}

A prediction-style plot is about predicting observations (or the ensemble mean of observations) for a given value of the focal predictor. For this, we want to use the classic curved confidence intervals (or prediction intervals). A prediction-style plot will generally focus on \emph{total} effects, and so will often be suited for a univariate fit. An effect-style plot, however, is an attempt to visualize the effect of a focal predictor. In this case, we suggest the narrower confidence intervals. Here, it makes equal sense to think about the total effect of our focal predictor, from a univariate model, or the direct effects – after controlling for whatever – from a multivariate model. See Fig~\ref{fig:justify_ci_plots}.

\subsection{Predictions vs. effects}

We would like to summarize by talking about predictions vs. effects. Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor. The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only and depend on the anchor.

>>>>

Ordinary least squares (OLS) regression provides an easy way to express multivariate relationships as an additive linear combination of many input variables with the outcome. The coefficient estimates from OLS regression models are generally easy to interpret as the expected change in the outcome with a unit change in the continuous variable or as the expected outcome difference between the levels of the categorical variable with the baseline category. However, with additional complexities such as interactions between input variables, the resulting coefficients risk misinterpretation or interpretation complexity \citep{brambor_understanding_2006, berry_improving_2012}.

Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including nonlinear relationships between response and input variables e.g., interactions between and nonlinear transformation on the input variables, (and via splines for example), and nonlinear transformations via link functions. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLM models involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore \emph{predictions} and \emph{effects} \citep{fox2009effect, leeper2017package, lenth2018package}.  For example, as a way to plan, public health officials may want to know the effects of household \code{income}, \code{wealthindex}, etc, on the predicted probability that slum dwellers will have improved water services.

Both predictions and effects tell us what we expect the response to be for a given value of \emph{focal} input variable (scientific variable of interest). The major difference between the two lies in the construction of the associated confidence intervals. In particular, for predictions, we want to use the classic curved confidence intervals (or prediction intervals) which capture the uncertainty in the whole model and generally focus on the \emph{total} effects, and so will often be suited for a univariate model. On the other hand, effects, attempt to show the effect of a focal input variable. In this case, we suggest the narrower confidence intervals which focus on uncertainty in the effect of the focal input variable only, and so will often be suited for the direct effects – after controlling for other non-focal input variables – from a multivariate model.

When visually presented, predictions and effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the input variables whose estimates are usually, but not always, harder to interpret. Generating these quantities together with the associated confidence intervals for regression models has a number of challenges. In particular:
\begin{enumerate}
\item choice of representative values of focal input variable and the \emph{reference point} for non-focal input variables, especially in multivariate models
\item  uncertainty estimation -- appropriate choice of \emph{anchor} for computing confidence intervals; and how to incorporate the uncertainty due to non-focal input variables 
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)Ms)
\end{enumerate}

The most common way of choosing the representative values is taking unique levels of the focal input variable if discrete or taking appropriately sized quantiles if continuous, and then calculating the predictions while holding non-focal (other input variables other than the focal input variable) input variables at their reference point (values we chose for the non-focal input variables e.g., means) \citep{fox2009effect, hanmer2013behind}. This procedure generates -- predictions (at least in our case), \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to these quantities as predictions and effects since they should, for example, tell us what we would expect the response to be at a particular value or level of the input variable, for an ``average case''. More specifically, the basic idea is to compute the expected outcome by holding the non-focal input variables constant, for example at their mean, while varying the focal input variable, with the goal that estimates represent how the model responds to the changes in the focal input variable. As mentioned above, in our case, the key difference between predictions and effects is in the computation of the confidence intervals.

The most commonly used \proglang{R} software packages for generating predictions (\pkg{emmeans} and \pkg{effects}), by default, use the average of \emph{input} and \emph{predictor variables} as the reference point for continuous and categorical input variables, respectively. However, there are a number of choices one can make when considering this approach -- for example, in the presence of interaction, averaging the interactions (averaging the product of interacting predictor variables) versus taking the product of the averages of the interacting predictor variables (the default for \pkg{emmeans} and \pkg{effects}). For simple models (with no interactions), the two approaches give the same predictions. However, for complex models, we demonstrate that the two approaches can produce substantially different results and show that averaging the interactions closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal input variable:

@@

The horizontal blue, green and black lines are the mean predictions, i.e., $\bar{\hat{y}}$, the red ones are the means of the simulated $y$, i.e., $\bar{y}$, while the vertical dotted grey lines are the means of the corresponding focal input variable. The grey points are the binned simulated $y$. The trend lines represent the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other input variables at their average for the three methods and the true predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated value in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches closely match the true predictions. B: Even with the simple interaction between non-focal input variables, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and true predictions, respectively; in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not the proposed \pkg{varpred}.}

@@

In the absence of interaction (model~\ref{eq:simple_inter_higher_no_interaction}), the three approaches produce the same estimates, which match the simulated values, Fig~\ref{fig:justify_plots}A. However, in the presence of interaction, even as simple as the one in model~\ref{eq:simple_inter_higher}, the estimates starts to differ. In particular, \pkg{emmeans} and \pkg{effects} give similar estimates ($\bar{\hat{y}}$) but different from the \pkg{varpred}'s which, however, is very close to the simulated average ($\bar{y}$), Fig~\ref{fig:justify_plots}B. To generate Fig~\ref{fig:justify_plots}A, all the three packages average $x_2$ and $x_3$, i.e., the input variables. On other hand, to generate Fig~\ref{fig:justify_plots}B, the difference in the estimates lies on how each of the packages average the interaction term ($x_2x_3$). In particular \pkg{emmeans} and \pkg{effects} use input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} uses predictor variables to compute $\over{x_2x_3}$. In the case of Fig~\ref{fig:justify_plots}A, averaging the input variables is equivalent to averaging the predictor variables.

One may be interested in the uncertainties associated with the focal input variable of interest only, excluding other uncertainties due to other non-focal input variables -- effects. However, currently, the two packages do not provide straightforward way do achieve this. To illustrate this, we generate predictions of $x_2$ from model~\ref{eq:simple_inter_higher_no_interaction} together with the associated $95\%$ confidence bands, as shown in Fig~\ref{fig:justify_ci_plots}.

\begin{figure}[h]
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal input variable, i.e., the model center, corresponds to the effects of the focal input variable from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal input variables. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal input variable.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence bands are much wider because they include uncertainties associated with the intercept and non-focal input variables, but narrower and crosses at the mean of the focal input variable in \pkg{varpred}. We refer to this point as the \emph{center point} and the approach as \emph{mean-based}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal input variable we are more certain about, i.e., mean of the predictor variable corresponding to the focal input variable. For simple models, the point where the confidence bands crosses is the model center; it corresponds to anchor we choose, in this case, the mean of the focal input variable, but we generally choose the mean of the focal predictor (predictors corresponding to the focal input variable).

For linear models (including models with complex interactions), the averaging is done on the linear scale, i.e., linear averaging. As a result the bias in the expected (mean) predictions can easily be corrected by performing predictions from the \emph{model center} (average of the predictor variables). The mean prediction is the prediction at the model center. On the other hand, when dealing with GL(M)Ms with nonlinear link functions, the averaging is done on the nonlinear scale, i.e., nonlinear averaging which makes estimating correct predictions much harder. In particular, bias in expected mean prediction induced by nonlinear transformation of the response variable can lead to inaccurate predictions. One way to deal with this is to make predictions on the transformed scale (linear predictor scale), and then back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation, for example, second order Taylor approximation implemented in \pkg{emmeans} \citep{lenth2018package}. In linear models, mean-based gives estimates consistent with the observed values, however, in models with nonlinear link functions, this is not always the case. An alternative to mean-based reference point, i.e., averaging non-focal predictor variables, is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictor variables and then averaging across the values of the focal predictor variable \citep{hanmer2013behind}.

Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor. The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only and depend on the anchor.

@@

\item \textbf{Input variables}: Refers to the observed (or scientific) variables underlying an inference or exploration. For example, the regression models described by Equation~\ref{eq:simple_inter_higher_no_interaction} and Equation~\ref{eq:simple_inter_higher} both have $3$ input variables -- $x_1, x_2, x_3$.
\item \textbf{Focal and non-focal input variables}: Focal input variable refers to the input variable we are interested in. For multivariate models, all other remaining input variables are non-focal input variables. For example, consider model described by Equation~\ref{eq:simple_inter_higher_no_interaction}. Suppose we are interested in the effect of $x_1$ on the predicted $y$, then $x_1$ is the focal input variable while $x_2$ and $x_3$ are non-focal non-focal variables.
\item \textbf{Model matrix}: Refers to the design matrix whose rows include all combination of input variables appearing in the interaction terms. Consider an example for three hypothetical households -- the first household head is a Christian with an income of $\$ 50$, the second household head is Muslim with an income of $\$ 100$ while the third household head is a Jew with an income of $\$ 77$. Suppose we want to model the household size (\code{hhsize}) as function of these household characteristics, i.e., $$\mathrm{hhsize} = \beta_0 + \beta_1\times\mathrm{income} + \beta_{2[r]}\times\mathrm{religion} + \mathrm{error}.$$ The model matrix corresponding to this model is given by

@@

\end{bmatrix}.$$ The first column represents the constant term in our model, $\beta_0$. For continuous input variables, the representation in the model matrix is the same as the corresponding input variables (for example second column representing income). For categorical variables, however, by default, the model matrix  creates additional dummy variables using the reference cell parameterization. This means that, if an input variable has $L$ factor levels, then there will be $L-1$ dummy columns representing all but the first level created in the model matrix. In our example, the column \code{religion} is missing and instead we have \code{religionJew} and \code{religionMuslim}; the missing category \code{religionChristian} is treated as the reference category.
\item \textbf{Predictor variables (or simply predictors)}: Refer to the X-variables which go into the regression model; and represent columns in the model matrix. Each input variable may correspond to one or more predictor variables. In particular, variables with more than two categories (for instance religion in our previous example), or input variables modeled with interactions, a spline or polynomial interactions, will correspond to more than one predictor variables. For example, the regression models described by Equation~\ref{eq:simple_inter_higher_no_interaction} and Equation~\ref{eq:simple_inter_higher} have $4$ predictors -- $x_1, x_2, x_3$ plus intercept term and $5$ predictors -- $x_1, x_2, x_3, x_2x_3$ plus the intercept term, respectively.
\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more predictor variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of predictor variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots} (with interaction). In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for center point. If we have more than one predictor variables associated with the focal input variable, however, the model center does not correspond to any given value of the variable, and we don't get the nice crossing we get in simpler cases. See Fig~\ref{fig:pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal input variables, when estimating the predictions. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate CIs becomes important. This is not true for the more conventional prediction plots: we can choose any anchor, e.g., zero-anchored, but model center seems like a natural, stable choice, i.e., center-anchored. In Fig~\ref{fig:justify_anchors}, we compare effects plots anchored at the minimum and center (mean) of the focal input variable.

@@
